---
title: "Nutrition Programming"
author: "By Moctar Aboubacar"
date: "1/25/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T)
```

### *Optimizing the allocation of supercereal food commodities to flood-affected Wards in Nepal in 2019*

Annual flooding in Nepal's Terai region exacerbates an already alarming child malnutrition situation. The World Food Programme (WFP) crafts its response to floods based on an assessment of affected areas and persons (extent, severity and drivers), existing responses, intra-UN programming coordination, and available resources.

The standard emergency nutrition intervention in this case is known as a Blanket Supplementary Feeding Programme (BSFP). Through BSFP, WFP transfers vitamin and mineral-fortified 'supercereal' to all pregnant and lactating women, and children between the ages of 6-59 months, within a given geographical area. Household level targeting is not done for BSFP; all eligible persons within the geographical areas receive the assistance. In July 2019, Nepal's Terai region suffered widespread flooding, with at least 6 Districts severely affected.

The nature of BSFP distribution poses a unique challenge in the context of flooding however, as unlike emergencies which tend to affect areas uniformly (such as drought), flooding can be highly localized. Given this, a blanket food distribution is likely to include people who are not in need of the food--in this case, pregnant and lactating women, and children 6-59 months who were not directly affected by flooding (for example, whose household food stores were not destroyed, etc.). Given limited resources, WFP's assistance should maximize assistance to directly affected populations.

The goal of this analysis is to achieve that optimization across Wards (the lowest administrative unit in Nepal. Wards are nested into Palikas, nested into Districts. A Ward may have a population of between 2,000-8,000 inhabitants) in the 6 most flood-affected Districts. This analysis treats the problem as a linear programming constrained optimization problem, specifically as a knapsack problem in which the inclusion of a unit (here a Ward) contributes to an overal objective function (estimated number of directly affected people), but at a cost (the constraint) of resources (in metric tonnes of supercereal). This analysis optimizes the combination of Wards in which to intervene at different levels of supercereal metric tonnes (representing different scenarios/possible resourcing levels).

#### **Data**

We start by loading the data, which is **(description of data) + of LPSolve** 

```{r warning=FALSE, message=FALSE}
# load packages
require(lpSolve)
require(tidyverse)
require(directlabels)

# import data
dat <- read.csv("C:/Users/Leurre/Desktop/r stuff/BFSP LP/Ward Exposure rank v1.csv")
glimpse(dat)
```

A quick glimpse shows that the data is a little messy. Part of this is because we used Excel to get the base data, and part of it is due to a number of missing data for wards in which satellite coverage was not available for the inundation period. The first rows for Dhanusa district represent wards for which satellite data was unavailable, and we are unable to produce estimates of affected population here.

```{r eval = TRUE, results = 'hide'}
# eliminate excel N/As for nonapplicable wards
dat[,c(17:22)]<- as.data.frame(apply(dat[,c(17:22)], 2,
                                       function(x){
                                         as.numeric(gsub("#N/A", "", x))
                                         }))

# 77 NAs, leaving 313-77 = 236 wards under consideration
sum(is.na(dat$plw_affect_perc)) 

# all NAs are dropped from the analysis, except those from the nutrition assessment
dat <- na.omit(dat[,1:23]) 
```

The percentage of population affected (pop_affect_perc) is a measure of the estimated percentage of people affected by flooding in each ward. From this measure (obtained via overlaying settlement-level population estimates with Sentinel-1 inundation mapping) we obtain an estimate of the number of affected pregnant and lactating women and of children under 5 in each Ward. This parameter becomes the objective function to maximize in the models.


#### **Ward resource allocation as a knapsack problem**

text here...we test the idea on first just 180 MTs (maybe also print a summary to show which districts are most concerned)

```{r eval = T, results = 'hide'}
# translate beneficiaries into KGs
dat$kg_req <- (dat$children_tot_num*6) + (dat$plw_tot_num*12)


# Linear programming algorithm for knapsack problem
# objective function: the number of people affected
dat <- mutate(dat, plw_children_affect_tot = plw_affect_num + children_affect_num)
knapsack.obj <- dat$plw_children_affect_tot

# constraints
knapsack.con <- matrix(dat$kg_req, nrow = 1, byrow = TRUE)
knapsack.dir <- "<="
knapsack.rhs <- c(180000) # test for 180 MT first

# LP function first iteration
ksSolution <- lp("max", 
                 knapsack.obj, 
                 knapsack.con, 
                 knapsack.dir, 
                 knapsack.rhs, 
                 all.bin = TRUE)

solution <- ksSolution$solution

sum(solution) # 40 wards are selected with 180 MT, stored in the solution vector

```


#### **Calculation across multiple MT options**

the whole point of this is to have results across multiple possibilities of MT.
we treat MTs as a discrete value still, but simulate from 300 to 975 MTs (true?) to see what our coverage would be at different levels of funding and supplies

```{r results = 'hide'}
# we repeat the process, this time obtaining the wards selected under different scenarios (300-975 MT, by increment of 25 MT)
numbers <- c(300000:975000)

# progressing by intervals of 25 MTs
ration_kg <- numbers[seq(1, length(numbers), 25000)] 

# create empty vector for the wards
wards <- vector() 

# number of columns for data frame
colLength <- length(dat) 

for (i in seq_along(ration_kg)){
  
  a <- lp("max", 
          knapsack.obj, 
          knapsack.con, 
          knapsack.dir, 
          ration_kg[i], 
          all.bin = TRUE)
  
  wards[i] <- sum(a$solution)
  
  dat[colLength + i] <- a$solution
}

# a list of the number of wards covered by each MT amount
results_lp <- data.frame(wards, ration_kg) 
```

#### **Results**

text goes here: we have some questions to operationalize these results:
1. how many wards total are reachable under different MT scenarios?
2. How many wards per district are reachable under different MT scenarios?
3. What percentage of all affected people are covered by different MT levels?

##### *1. Total wards reachable*

text text text text goes here

```{r}
Wards_by_MT <- ggplot(results_lp, aes(x = (ration_kg/1000), y = wards))+
  geom_line(color = "indianred4", size = 1.2)+
  theme_bw()+
  labs(title = "Wards by possible BSFP MT Allocation",
       caption = "Out of a total 236 Wards",
       y = "Number of Wards",
       x = "BSFP Metric Tonnes")

Wards_by_MT
```



##### *2. Wards per district*

text here

```{r}
viztable1 <- dat %>% 
  group_by(District) %>% 
  summarise_at(vars(starts_with('V')), sum)

ration_mt <- ration_kg/1000
names(viztable1)[2:29] <- ration_mt # rename columns according to metric tonnage designation


viztable1 <- gather(viztable1, key = "MTs", value = "Wards",
                    c(2:29))
# viztable1$MTs <- as.integer(viztable1$MTs)


nominalDistricts <- c("grey60","grey60","grey60","grey60", "indianred4","grey60")

require(directlabels)

District_compare <- ggplot(viztable1, aes(x = MTs, y = Wards, group = District))+
  geom_line(aes(color = District), size = 1.1)+
  scale_color_manual(values = nominalDistricts)+
  scale_x_discrete(expand = c(0, 3))+
  geom_dl(aes(label = District, color = District), method = list(dl.combine('first.points', "last.points"), cex = 0.9))+
  theme(legend.position = "none")+
  labs(title = "By-District Ward Numbers per Metric Tonnage",
       subtitle = "Sarlahi consistently counts the most wards above 375 MTs",
       x = "BSFP Metric tonnes",
       y = "Number of Wards")+
  theme_bw()

District_compare
```

~~In the case of Sarlahi, this is driven by both smaller Ward sizes and higher proportions of affected persons per Ward.~~

##### *3. % affected people*

text goes here.
As expected the efficiency gains in delivery slow down at higher metric tonnes, suggesting that the analysis is doing what it should be; at higher levels of metric tonnes we start to cover more and more ''inefficient' Wards with fewer total population directly affected within them.

A humanitarian approach will seldom select the most efficient point of

```{r}
viztable2 <- mutate(dat, plw_children_affect_perc = plw_children_affect_tot/sum(plw_children_affect_tot)) 

names(viztable2)[26:53] <- ration_mt

viztable2 <- gather(viztable2, key = "MTs", value = "Wards",
                    c(26:53))

viztable2 <- viztable2 %>% 
  group_by(MTs, Wards) %>% 
  summarise(perc_affected = sum(plw_children_affect_perc)) %>% 
  filter(Wards == 1)

viztable2$MTs <- as.integer(viztable2$MTs)

Percentage_coverage <- ggplot(viztable2, aes(x = MTs, y = perc_affected))+
  geom_line(color = "indianred4", size = 1.2)+
  theme_bw()+
  labs(title = "Percentage Affected Coverage",
       subtitle = "500 MTs covers over 75% of affected people",
       x = "BSFP Metric tonnes",
       y = "% affected people reached")

Percentage_coverage
```

#### **Discussion**
- this won't be followed ward by ward, for funding reasons + political reasons
- but it can be a good guide
- more importantly, this kind of approach, supplemented by better vulnerability information (here we are just using exposure), including estimations of malnutrition at War level, could yield something better than this.
